### 2020-01-31 Common/chapters/deep_learning.tex
Johan and Julien: paper reproduction chapter continued

\chapter{Deep learning}
\label{ch:deep_learning}

\section{Introduction to deep learning}
Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence. Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them. In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way. There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc. What mainly differentiates deep learning from these algorithms is the concept of neural networks (see section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
Neural networks are inspired from the biological neural networks of the brain. These systems try to learn how to solve a problem based on the data they receive as input. Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics, etc. The presence of deep learning techniques in these use cases clearly attests that this technology is a breakthrough. Furthermore, since this field has recently gained interest (see section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

>>>>

\chapter{Deep learning}
\label{ch:deep_learning}
This chapter provides the theoretical foundation in deep learning for the understanding of the rest of the work. It starts with the historical story of the deep learning before describing what is a neural network. Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation, is explained. This part is followed with another devoted to a special type of neural networks, the "convolutional neural networks". They are used in many computer vision problems due to their great performance for these tasks. Finally, the concept of "transfer learning" is discussed since the entire chapter \ref{ch:transfer_learning} relies on it.
\section{Introduction to deep learning}
Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence. Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them. In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way. There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc. What mainly differentiates deep learning from these algorithms is the concept of neural networks (see section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
Neural networks are inspired from the biological neural networks of the brain. These systems try to learn how to solve a problem based on the data they receive as input. Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics, etc. The presence of deep learning techniques in these use cases clearly testifies to the enthusiasm of many areas for this technology. Furthermore, since this field has recently gained interest (see section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

### 2020-01-31 Common/chapters/paper_reproduction.tex
Johan and Julien: paper reproduction chapter continued

\chapter{Paper reproduction}
\label{ch:paper_reproduction}
This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI"\cite{07} presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a new transfer learning method as a way to improve the classification using other body parts and to overcome the lack of available data(see \ref{ch:transfer_learning}).\\

Furthermore, since the article does not provide the results their model got on the PROSTATEx challenge, the reproduction of the experiment can act like a benchmark to evaluate the generalization of the model to unseen data.

\section{Process overview}
Schematically, the entire experiment process can be represented as shown in figure \ref{fig:paper_reproduction_process}. The PROSTATEx dataset is composed of samples whose clinical significance is provided (labeled data) and samples without this information (unlabeled data). The latter are used for the challenge, which consists in predicting the probability of patients lesions to be malignant (class 1) or benign (class 0). In both cases, the data is processed before been used (see \ref{prostatex_data_processing} for details). After the processing, the labeled data are split into training, validation and test sets using respectively 80\%, 10\% and 10\% of all available data for each set. The training set is then given in batches as input to the neural network, which update its weight consequently. At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved.  When the training phase reaches the defined number of epochs, the model is then tested on the test set. This same model is also used to take part into the PROSTATEx challenge in order to output predictions into a .csv file for each patient and findings.

>>>>

\chapter{Paper reproduction}
\label{ch:paper_reproduction}
This chapter relies on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts, which overcome the lack of available data (see chapter \ref{ch:transfer_learning}).\\
Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest available dataset for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). This paper explains all the steps to follow with the goal of building a computer-aided-diagnosis (CAD) system for prostate cancer detection. Moreover, it also provides results about the algorithm performance, which can be used as a benchmark to compare with the results of the reproduction of the experiment. However, since the article gives no information about the results their model got on the official PROSTATEx challenge test set, the reproduction of the experiment will fill that gap.\\
This chapter is built top-down: first a superficial overview of the entire process is established in order to understand the purpose of the experiment as a whole. Then, each part or theoretical notion of the experiment is deeply described. The structure of the dataset is detailed, all steps of its processing are explained and the techniques used as verification of the proper functioning of the algorithm are presented. After the processing of the data, the training phase is deepened and all hyperparameters, options and implementation choices are given to ensure the reproductibility of the experiment. This part is followed by the presentation of the raw results, which is itself followed by their analysis in the "discussion" section.

\section{Process overview}
Schematically, the entire experiment process can be represented as shown in figure \ref{fig:paper_reproduction_process}. The PROSTATEx dataset is composed of samples whose clinical significance is provided (labeled data) and samples without this information (unlabeled data). The latter are used for the challenge, which consists in predicting the probability of patients lesions to be malignant (class 1) or benign (class 0). In both cases, the data is processed before been used (see \ref{prostatex_data_processing} for details). After the processing, the labeled data are split into training, validation and test sets using respectively 80\%, 10\% and 10\% of all available data for each set. The training set is then given in batches as input to the neural network, which update its weight consequently. At the end of each epoch, the current model is tested on the validation set (that the model has never seen previously) and the metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). If the current model has a bigger AUC and a bigger accuracy than the previous best model, it becomes the best model and it is saved.  When the training phase reaches the defined number of epochs, the model is then tested on the test set. This same model is also used to take part in the PROSTATEx challenge to output predictions into a .csv file for each patient and findings.
